Simulating Benefits, Costs, and Tradeoffs of Spatial Management in Marine Social-Ecological Systems

### 2023-10-25 03_marlin_paper.qmd
first look edits

# rdatas <- results_files[str_detect(results_files,'.Rdata')]
# 
# walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv)) # load results into global environment

@@

Bio-economic modeling, spatio-temporal modeling, marine protected areas,
fisheries management, marine spatial planning, conservation planning

@@

1.  Introduction
2.  Methods
    1.  Model Summary
    2.  Fleet Dynamics
    3.  Population Model
    4.  Case Studies
3.  Results
    1.  Coral Reefs
    2.  Pelagic Systems
4.  Discussion
    1.  Insights from Case Studies
    2.  Putting `marlin` in Context
    3.  General Recommendations for Use
5.  Acknowledgements
6.  Data Availability Statement
7.  References

@@

To illustrate, policies such as MPAs designed for both conservation and food
production must consider factors such as the optimal size and placement of a
protected area network given a wide array of life history, species
distributions, exploitation levels, fleet dynamics, and policy-dependent
behavior [@reimer2017], all of which may change over time, particularly given
the impacts of climate change. Efforts to effectively design spatial management
strategies such as MPAs are further constrained by a lack of empirical evidence
describing the size [@ovando2021a] and time required [@nickols2019] for MPAs to
produce substantial effects across a range of social-ecological settings
[@mcclanahan2021].

Many models have been developed to support the theory and design of spatial
management strategies (see @fulton2015 for a review of existing methods).
However, these models have tended to be either highly complex and tactical
models designed for use in a specific location, or to be extremely stylized
representations intended to provide theoretical insights with less relevance for
specific applications. This lack of accessible models suited for more tactical
application presents a challenge to stakeholders charged with marine spatial
planning exercises. It also presents a barrier to the scientific community,
wherein comparisons across spatial management simulation studies are clouded by
discrepancies in underlying model structures beyond differences in the core
phenomenon in question.

To help address this challenge, we present a novel bio-economic modeling tool
called [`marlin`](https://danovando.github.io/marlin/) that allows for efficient
simulation of spatio-temporal biological and economic dynamics. `marlin` allows
users to simulate the impacts of marine management policies across a range of
species targeted by various fleets across heterogeneous and dynamic seascapes.
This model can help users seek Pareto-optimal solutions that produce win-wins or
minimize trade-offs across multiple management objectives, such as food
production and conservation [@lester2013].

@@

These species can be caught by fishing fleets. A fishing fleet is defined in
`marlin` by a set of fishers that have the same fishing skill, prices, and
contact selectivity [@sampson2014] for individual species (each of which we
denote as a *metier*). For example, both a longline and purse seine fleet may
capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline fleet can
be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
It is important to note that contact selectivity reflects the ability of the
fishing method in question to capture fish of different sizes that come into
"contact" with the gear. The contact selectivity of each of the specified
*metiers* will then interact with the distribution of fish sizes and fishing
effort in space simulated by `marlin` to produce a net "population" selectivity,
which may differ from the individual contact selectivities of each of the
*metiers* [@waterhouse2014; @sampson2014]. Each fishing fleet distributes its
fishing effort in space according to a specified spatial allocation rule (see
*Spatial Allocation of Effort* section), for example in proportion to
profit-per-unit-effort, conditional on management policies in places such as
quotas and/or the presence of any spatial restrictions such as MPAs.

@@

To our knowledge, MPAs are never randomly placed; their locations are decided
based on stakeholder negotiation given available information and objectives. In
theory, the design of MPA networks can be optimized through the use of a
modeling framework, and depending on the validity of the model, this process is
likely to produce better outcomes than manually-designed strategies
[@rassweiler2012; @rassweiler2014]. However, designing optimal MPA networks
becomes increasingly difficult as the range of objectives and the complexity of
the model increase. Therefore, while `marlin` provides some support for
[optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html),
we focus here on the design and performance of these more rule of thumb design
strategies that may be more accessible to a wider range of users. We allow all
MPAs to be designed in a mosaic fashion in these examples, but users can easily
extend the analysis to compare outcomes between contiguous (MPA is made up of
one continuously connected block) and mosaic (MPAs can be separated in space)
MPA designs [@pons2022] by developing and passing their own MPA design
algorithms to the model.

@@

```{r}
#| label: fig-coral
#| fig-cap: "Distribution of fish spawning biomass (cell color, yellow = higher, blue = lower) each season (columns) under unfished conditions. The x-axis represents longitude, y-axis latitude. Numbers show the approximate location of each fleet's (one or two) port. Columns indicate quarterly seasons that repeat each year."

 patch_biomass %>%
  group_by(critter, step) %>%
  mutate(sbiomass = biomass / max(biomass)) %>%
  ungroup() %>%
  filter(between(step,1,2)) %>%
  mutate(step = fct_recode(as.factor(step), "2" = "1.25", "3" = "1.5", "4" = "1.75")) %>% 
  mutate(Season = step) %>% 
  ggplot()+
  geom_tile(aes(x,y, fill = sbiomass),show.legend = FALSE) +
  geom_text_repel(data = coral_ports, aes(x, y, label = fleet), color = "red") +
  facet_grid(critter~Season, labeller = labeller(critter = titler, Season = label_both)) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_viridis_c() + 
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank())

@@

We model our pelagic case study loosely on the characteristics of the Western
and Central Pacific Ocean (WCPO) tuna fisheries. We simulate trajectories of
`r n_distinct(names(blue_fauna))` species commonly caught in the region,
including both the highly abundant skipjack tuna and the heavily depleted
oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb). We use
publicly available data on catch-per-unit-effort of each of these species from
the WCPO as a very rough proxy for baseline habitat distributions, noting that
where possible, fishery-independent abundance indices would be preferable
(@fig-blue).

These pelagic species are caught by a longline fleet that primarily targets
large adult tunas such as bigeye and yellowfin for high-grade consumption, and a
purse seine fishery that primarily targets skipjack tunas for bulk canning.
Contact selectivities were modeled as logistic for the longline fleet, and
dome-shaped for the purse seine fleet, with selectivity values set to broadly
match the patterns reported in relevant [stock
assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
for each species. Both fleets operate under open-access dynamics with an effort
cap. The effort cap was set at the level of effort that resulted in the desired
levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue),
intended to simulate a scenario where managers step in to prevent further
expansion of fishing effort in a fully developed fishery. For
forward-simulation, open-access dynamics can result in effort decreasing in
response to profitability, but cannot result in effort beyond the effort cap set
for each fleet.

@@

```{r}
#| label: fig-blue-ssb
#| fig-cap: "Spawning biomass divided by unfished spawning biomass in the time period prior to implementation of MPAs for the pelagic case study."

critter_biomass %>%
  ungroup() %>% 
  left_join(sci_to_com, by = "critter") %>%
  select(-critter) %>%
  rename(critter = common_name) %>%
  filter(step == max(step)) %>% 
  ggplot(aes(reorder(critter, depletion), depletion)) +
  geom_col() +
  scale_y_continuous(limits = c(0, 1), name = "SSB/SSB0", expand = expansion(mult = c(0.0,0.05))) +
  labs(x = "") + 
  coord_flip()

range_shift_plot <- sq_habitat | rs_habitat

@@

coral_frontier <- tmp %>%
  filter(prop_mpa < 0.7) |>
  ggplot(aes(delta_biodiv, delta_yield)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_path(alpha = 0.8, aes(color = placement_strategy),size = 1.1) +
  geom_point(aes(fill = prop_mpa), shape = 21, size = 3) +
  # geom_text_repel(data = frontier_labels, aes(label = pmpa), min.segment.length = 0, box.padding = 0.5) +
  scale_x_continuous(name = "Change in Total SSB/SSB0", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2), limits = c(0,0.3)) +
  scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) +
  scale_color_manual(values = c("tomato", "steelblue"),name = '') +
  # guides(fill = guide_legend(order = 1, override.aes = list(size = 4))) +
  scale_fill_viridis(name = "% MPA", labels = scales::label_percent(accuracy = 1),guide = guide_colorbar(ticks.colour = "black",frame.colour = "black"), option = "cividis") +
  labs(tag = "C")  +
  theme(
    legend.position = c(0.73, .73),
    legend.background = element_rect(fill = "transparent"),
    legend.text = element_text(size = 8)
  )
coral_fleet_yield <- coral_fleet_results |> 
  mutate(fleet= fct_relabel(fleet, titler)) %>% 
  ggplot(aes(prop_mpa, delta_yield, color = fleet)) +
    geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line() +
  facet_wrap( ~ placement_strategy, labeller = labeller(placement_strategy = titler)) +
  scale_x_continuous(name = "MPA Size",
                     labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(name = "Yield",
                     labels = scales::label_percent(accuracy = 1)) + 
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 8),
        legend.margin = margin())  + 
    scale_color_discrete(name = '') + 
    labs(tag = "A")

((coral_fleet_yield / coral_critter_bio) | (coral_frontier + plot_layout(guides = "keep")) ) & theme(strip.text = element_text(size  = 6), plot.margin = margin(.05, .05, .05, .05, "cm"), panel.spacing = unit(0.25, "lines"), legend.text = element_text(size = 6))

@@

@@

# blue_frontier / blue_critter_bio + plot_layout(heights = c(1,2))
(blue_fleet_yield / 
blue_critter_bio) + plot_layout(heights = c(1.5,2))
```

Our coastal coral reef case study results show the sensitivity of conservation
and food security outcomes of MPAs to alternative design scenarios, holding
constant other social-ecological variables. The *Spawning Ground* MPA network
was able to provide more equitable yield outcomes across the two fishing
communities, whereas the *Target Fishing* MPA network only benefited the more
industrial Fleet Two (@fig-coral-results A). This is because Fleet One has two
primary fishing grounds; the spawning grounds, and the offshore area where the
deep snapper live. Fleet Two primarily fishes in the northern portions of the
simulation grid. Under the *Spawning Ground* strategy, while Fleet One quickly
losses fishing grounds on the spawning grounds, it can compensate for this by
moving offshore and fishing the deep snapper population harder, while also being
able to fish some of the spillover of snapper from the spawning ground closure.
This displacement of Fleet One's fishing effort is why biomass of deep snapper
actually declines under the *Spawning Ground* strategy.

By protecting both the spawning ground and the offshore areas, the *Target
Fishing* network provides conservation benefits to all of the species, in
contrast to the *Spawning Ground* strategy that only provides meaningful
conservation gains to the snapper population up until very large MPA sizes. Our
result that protection of a dedicated spawning ground did not provide
substantially greater conservation outcomes for the species using that spawning
ground (groupers and sharks) than an alternative design strategy is supported by
other modeling studies that show that the impacts of spawning ground protection
on conservation and yields may be highly variable, and that displacement of high
levels fishing effort from the spawning grounds can offset potential
conservation gains of the protection [@grüss2014]. That being said there is also
evidence for the benefits of spawning aggregation protection [@erisman2015] ,
and our results are not nearly sufficiently resolved to provide any general
statements as to the relative value of spawning aggregation protection relative
to other design strategies.

Studies such as @gilman2020, @abbott2012, @davies2018 and @pons2022 provide
empirical support for our case study results showing that MPAs can produce
trade-offs across the conservation of different species and fleets when the
spatial-temporal distributions and life history traits of affected species are
heterogeneous, and MPAs result in some degree of effort displacement or
concentration in the remaining fishing grounds. @hampton2023 supports our
pelagic case study result showing that both conservation and fishery impacts of
MPAs on the highly mobile species of the open oceans can be limited unless
protected areas are very large. @brown2018 supports our result showing how
heterogeneity in the behavior and objectives of fishing fleets sharing an
ecosystem can affect the magnitude and equity of fishery reform outcomes.
@davies2017 discusses the importance of considering the potential of climate
driven range shifts when assessing spatial management policies; our work builds
on this by allowing users to not just simulate species distributions but also
distribution of biomass and age composition of fish space and time under climate
change.

`marlin` does not represent trophic interactions. Anthropgenic changes in
species abundance can result in trophic cascades throughout affected food webs.
However, studies such as @gilman2020, @ovando2021a, @bruno2019 , and
@malakhoff2021 found no clear signs of MPA driven trophic cascades within the
first decades of protection. Signals of trophic cascades may be masked by
variations in the direction and strength of species interactions driven by
environmental context [@liu2022], or may simply take longer to develop
detectable effects than the coverage of many time series of MPAs. While `marlin`
does not incorporate trophic interactions, what empirical evidence we have does
not suggest that management-mediated trophic cascades are so common and clear
that they must be incorporated into any credible multi-species simulation model.
There is value in capturing other dynamic bio-economic processes, as done here,
even if that comes at the expense of modeling trophic interactions. However,
research on the trophic impacts of spatial-temporal management actions is
clearly of value and models like Atlantis [@audzijonyte2019], Ecopath with
Ecosim [@christensen2004], and EASI-Fish [@griffiths2019] can help users explore
those kinds of trophic processes, in the manner of @baskett2007.

Marine social-ecological systems are highly complex. The model presented here is
designed to help users explore the impacts of different variables on policy
outcomes. But, that freedom comes at a cost, namely that users have a large
number of options at their disposal that they must decide on. Our recommended
strategy is for users to try and narrow down a list of parameters that they feel
are sufficiently "known" and another list of parameters that are highly
uncertain and / or contentious which the users feel may impact results. Where
possible, parameters from locally estimated stock assessment models can be used
to provide a foundation around which sensitivity analyses around specific
uncertainties of interest [@berger2017].

Given a set of model runs, the choice of whether to use those results
"strategically" or "tactically" will depend on the needs of the user. That being
said, since. `marlin` is not fit to data directly it is more easily applied to
strategic questions. For example, the two case studies presented in this paper
provide general strategic insights into the sensitivity of outcomes to design
choices range shifts. The extent to which users are comfortable using the
outputs of `marlin` tactically will depend on the confidence they have in their
parameterization of the model runs relative to the precision policy-makers
require in order to make a decision.

Even the most complex marine model is a stylized cartoon of the true dynamics of
ocean ecosystems. However, for all their limitations, models such as `marlin`
can help users understand factors that drive the performance of marine
management strategies. The goal of this tool is to empower people to design
policies based on evaluation of key uncertainties and trade-offs, and in doing
so support more effective and equitable marine resource management.

>>>>

small_width <- 80 # figure width in mm

big_width <- 180 # width in mm

@@

bio-economic modeling
conservation planning
fisheries
marine protected areas
marine spatial planning
spatio-temporal modeling

@@

1.  Introduction
2.  Methods
    1.  Model Summary
    2.  Fleet Dynamics
    3.  Population Model
    4.  Case Studies
3.  Results
    1.  Coral Reefs
    2.  Pelagic Systems
4.  Discussion
    1.  Insights from Case Studies
    2.  Putting `marlin` in Context
    3.  General Recommendations for Use
5.  Acknowledgements
6.  Data Availability Statement
7.  References
8. Figure Legends

@@

To illustrate, policies such as MPAs designed for both conservation and food
production must consider factors such as the optimal size and placement of a
protected area network given a wide array of life history, species
distributions, exploitation levels, fleet dynamics, and policy-dependent
behavior [@reimer2017], all of which may fluctuate over time, particularly given
the impacts of climate change. Efforts to effectively design spatial management
strategies such as MPAs are further constrained by a lack of empirical evidence
describing the size [@ovando2021a] and time required [@nickols2019] for MPAs to
produce substantial effects across a range of social-ecological settings
[@mcclanahan2021].

Many models have been developed to support the theory and design of spatial
management strategies (see @fulton2015 for a review).
However, these models have tended to be either highly complex and tactical
models designed for use in a specific location, or to be extremely stylized
representations intended to provide theoretical insights with less relevance for
specific applications. This lack of accessible models capable of representing reasonable amounts of complexity presents a challenge to stakeholders charged with marine spatial
planning exercises. It also presents a barrier to the scientific community,
wherein comparisons across spatial management simulation studies are clouded by
discrepancies in underlying model structures beyond differences in the core
phenomenon in question.

To help address this challenge, we present a bio-economic modeling tool
called [`marlin`](https://danovando.github.io/marlin/) that allows for efficient
simulation of spatio-temporal biological and economic dynamics. `marlin` allows
users to simulate the impacts of marine management policies across a range of
species targeted by various fleets across heterogeneous and dynamic seascapes.
This model can help users seek Pareto-optimal solutions that produce win-wins or
minimize trade-offs across multiple management objectives, such as food
production and conservation [@lester2013].

@@

These species can be caught by fishing fleets. A fishing fleet is defined in
`marlin` by a set of fishers that have the same fishing skill, prices, and
contact selectivity [@sampson2014] for individual species (each of which we
denote as a *metier*). For example, both a longline and purse-seine fleet may
capture bigeye (*Thunnus obesus*) and skipjack (*Katsuwonus pelamis*) tunas, but the longline fleet can
be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
It is important to note that contact selectivity reflects the ability of the
fishing method in question to capture fish of different sizes that come into
"contact" with the gear. The contact selectivity of each of the specified
*metiers* will then interact with the distribution of fish sizes and fishing
effort in space simulated by `marlin` to produce a net "population" selectivity,
which may differ from the individual contact selectivities of each of the
*metiers* [@waterhouse2014; @sampson2014]. Each fishing fleet distributes its
fishing effort in space according to a specified spatial allocation rule (see
*Spatial Allocation of Effort* section), for example in proportion to
profit-per-unit-effort, conditional on management policies in places such as
quotas and/or the presence of any spatial restrictions such as MPAs.

@@

To our knowledge, MPAs are never randomly placed; their locations are decided
based on stakeholder negotiation given available information and objectives. In
theory, the design of MPA networks can be optimized through the use of a
modeling framework, and depending on the validity of the model, this process is
likely to produce better outcomes than manually-designed strategies
[@rassweiler2012; @rassweiler2014]. However, designing optimal MPA networks
becomes increasingly difficult as the range of objectives and the complexity of
the model increase. Therefore, we focus here on the design and performance of these more rule of thumb design
strategies that may be more accessible to a wider range of users. We allow all
MPAs to be designed in a mosaic fashion in these examples, but users can easily
extend the analysis to compare outcomes between contiguous (MPA is made up of
one continuously connected block) and mosaic (MPAs can be separated in space)
MPA designs [@pons2022] by developing and passing their own MPA design
algorithms to the model.

@@

@@

We model our pelagic case study loosely on the characteristics of the Western
and Central Pacific Ocean (WCPO) tuna fisheries. Note that this is an illustrative example only and simulated stock status, species distributions, or outcome projections presented here should not be interpreted as a indicative of the current or future state of the WCPO. We simulate trajectories of
`r n_distinct(names(blue_fauna))` species commonly caught in the region,
including both the highly abundant skipjack tuna and the heavily depleted
oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb). We use
publicly available data on catch-per-unit-effort of each of these species from
the WCPO as a very rough proxy for baseline habitat distributions, noting that
where possible, fishery-independent abundance indices would be preferable
(@fig-blue).

These pelagic species are caught by a longline fleet that primarily targets
large adult tunas such as bigeye and yellowfin for high-grade consumption, and a
purse-seine fishery that primarily targets skipjack tunas for bulk canning.
Contact selectivities were modeled as logistic for the longline fleet, and
dome-shaped for the purse-seine fleet, with selectivity values set to broadly
match the patterns reported in relevant [stock
assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
for each species. Both fleets operate under open-access dynamics with an effort
cap. The effort cap was set at the level of effort that resulted in the desired
levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue),
intended to simulate a scenario where managers step in to prevent further
expansion of fishing effort in a fully developed fishery. For
forward-simulation, open-access dynamics can result in effort decreasing in
response to profitability, but cannot result in effort beyond the effort cap set
for each fleet.

@@

@@

@@

Our coastal coral reef case study results show the sensitivity of conservation
and food security outcomes of MPAs to alternative design scenarios, holding
constant other social-ecological variables. The *Spawning Ground* MPA network
was able to provide more equitable yield outcomes across the two fishing
communities, whereas the *Target Fishing* MPA network only benefited Fleet Two (@fig-coral-results A). This is because Fleet One has two
primary fishing grounds; the spawning grounds, and the offshore area where the
deep snapper live. Fleet Two primarily fishes in the northern portions of the
simulation grid. Under the *Spawning Ground* strategy, while Fleet One quickly
losses fishing grounds on the spawning grounds, it can compensate for this by
moving offshore and fishing the deep snapper population harder, while also being
able to fish some of the spillover of snapper from the spawning ground closure.
This displacement of Fleet One's fishing effort is why biomass of deep snapper
actually declines under the *Spawning Ground* strategy.

By protecting both the spawning ground and the offshore areas, the *Target
Fishing* network provides conservation benefits to all of the species, in
contrast to the *Spawning Ground* strategy that only provides meaningful
conservation gains to the snapper population up until very large MPA sizes. Our
result that protection of a dedicated spawning ground did not provide
substantially greater conservation outcomes for the species using that spawning
ground (groupers and sharks) than an alternative design strategy is supported by
other modeling studies that show that the impacts of spawning ground protection
on conservation and yields may be highly variable, and that displacement of high
levels fishing effort from the spawning grounds can offset potential
conservation gains of the protection [@grüss2014]. That being said there is also
evidence for the benefits of spawning aggregation protection [@erisman2015] ,
and our results are not nearly sufficiently resolved to provide any general
statements as to the relative value of spawning aggregation protection relative
to other design strategies. Further research could for example alter both the distribution of species in space and the susceptibility of the species to fishing gear when aggregated.

Studies such as @gilman2020, @abbott2012, @davies2018 and @pons2022 provide
empirical support for our case study results showing that MPAs can produce
trade-offs across the conservation of different species and fleets when the
spatial-temporal distributions and life history traits of affected species are
heterogeneous, and MPAs result in some degree of effort displacement or
concentration in the remaining fishing grounds. @hampton2023 supports our
pelagic case study result showing that both conservation and fishery impacts of
MPAs on the highly mobile species of the open oceans can be limited unless
protected areas are very large. @brown2018 supports our result showing how
heterogeneity in the behavior and objectives of fishing fleets sharing an
ecosystem can affect the magnitude and equity of fishery reform outcomes.
@davies2017 discusses the importance of considering the potential of climate
driven range shifts when assessing spatial management policies; our work builds
on this by allowing users to not just simulate species distributions but also
distribution of biomass and age composition of fish in space and time under climate
change.

`marlin` does not represent trophic interactions. Anthropgenic changes in
species abundance can result in trophic cascades.
However, studies such as @gilman2020, @ovando2021a, @bruno2019 , and
@malakhoff2021 found no clear signs of MPA driven trophic cascades within the
first decades of protection. Signals of trophic cascades may be masked by
variations in the direction and strength of species interactions driven by
environmental context [@liu2022], or may simply take longer to develop
detectable effects than the coverage of many time series of MPAs. While `marlin`
does not incorporate trophic interactions, what empirical evidence we have does
not suggest that management-mediated trophic cascades are so common and clear
that they must be incorporated into any credible multi-species simulation model.
There is value in capturing other dynamic bio-economic processes, as done here,
even if that comes at the expense of modeling trophic interactions. However,
research on the trophic impacts of spatial-temporal management actions is
clearly of value and models like Atlantis [@audzijonyte2019], Ecopath with
Ecosim [@christensen2004], and EASI-Fish [@griffiths2019] can help users explore
those kinds of trophic processes, in the manner of @baskett2007.

Marine social-ecological systems are highly complex. The model presented here is
designed to help users explore the impacts of different variables on policy
outcomes. But, that freedom comes at a cost, namely that users have a large
number of options at their disposal that they must decide on. Our recommended
strategy is for users to try and narrow down a list of parameters that they feel
are sufficiently "known" and another list of parameters that are highly
uncertain and / or contentious which the users feel may impact results. Where
possible, parameters from locally estimated stock assessment models can be used
to provide a foundation around which sensitivity analyses around specific
uncertainties of interest [@berger2017], though care should be taken interpreting population selectivity curves estimated from stock assessment as contact selectivity curves required as inputs to `marlin` [@sampson2014].

Given a set of model runs, the choice of whether to use those results
"strategically" or "tactically" will depend on the needs of the user. That being
said, since. `marlin` is not fit to data directly it is more easily applied to
strategic questions. For example, the two case studies presented in this paper
provide general strategic insights into the sensitivity of outcomes to design
choices range shifts. The extent to which users are comfortable using the
outputs of `marlin` tactically will depend on the confidence they have in their
parameterization of the model relative to the precision policy-makers
require in order to make a decision.

Even the most complex marine model is a stylized cartoon of the true dynamics of
ocean ecosystems. However, for all their limitations, models
can help users understand factors that drive the performance of marine
management strategies. The goal of this tool is to empower people to design
policies based on evaluation of key uncertainties and trade-offs, and in doing
so support more effective and equitable marine resource management.

```{r}
#| label: fig-coral
#| fig-cap: "Distribution of fish spawning biomass (cell color, yellow = higher, blue = lower) each season (columns) under unfished conditions. The x-axis represents longitude, y-axis latitude. Numbers show the approximate location of each fleet's (one or two) port. Columns indicate quarterly seasons that repeat each year. "

 fig_1 <- patch_biomass %>%
  group_by(critter, step) %>%
  mutate(sbiomass = biomass / max(biomass)) %>%
  ungroup() %>%
  filter(between(step,1,2)) %>%
  mutate(step = fct_recode(as.factor(step), "2" = "1.25", "3" = "1.5", "4" = "1.75")) %>% 
  mutate(Season = step) %>% 
  ggplot()+
  geom_tile(aes(x,y, fill = sbiomass),show.legend = FALSE) +
  geom_text_repel(data = coral_ports, aes(x, y, label = fleet), color = "red") +
  facet_grid(critter~Season, labeller = labeller(critter = titler, Season = label_both)) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_viridis_c() + 
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank())

 ggsave(
   filename = file.path(results_path, "fig_1.pdf"),
   fig_1,
   width = small_width,
   units = "mm"
 )

 fig_1

```{r}
#| label: fig-blue-ssb
#| fig-cap: "Spawning biomass divided by unfished spawning biomass in the time period prior to implementation of MPAs for the pelagic case study. Simulated species include skipjack tuna, yellowfin tuna (*Thunnus albacares*), bigeye tuna, shortfin mako (*Isurus oxyrinchus*), swordfish (*Xiphias gladius*), albacore tuna (*Thunnus alalunga*), blue shark (*Prionace glauca*), silky shark (*Carcharhinus falciformis*), and oceanic whitetip shark (*Carcharhinus longimanus*)."

fig_2 <- critter_biomass %>%
  ungroup() %>% 
  left_join(sci_to_com, by = "critter") %>%
  select(-critter) %>%
  rename(critter = common_name) %>%
  filter(step == max(step)) %>% 
  ggplot(aes(reorder(critter, depletion), depletion)) +
  geom_col() +
  scale_y_continuous(limits = c(0, 1), name = "SSB/SSB0", expand = expansion(mult = c(0.0,0.05))) +
  labs(x = "") + 
  coord_flip()

 ggsave(
   filename = file.path(results_path, "fig_2.pdf"),
   fig_2,
   width = small_width,
   units = "mm"
 )

fig_2

range_shift_plot <- fit_3 <- sq_habitat | rs_habitat

 ggsave(
   filename = file.path(results_path, "fit_3.pdf"),
   fit_3,
   width = big_width,
   units = "mm"
 )

coral_frontier <- tmp %>%
  filter(prop_mpa < 0.7) |>
  ggplot(aes(delta_biodiv, delta_yield)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_path(alpha = 0.8, aes(color = placement_strategy),size = 1.1) +
  geom_point(aes(fill = prop_mpa), shape = 21, size = 3) +
  # geom_text_repel(data = frontier_labels, aes(label = pmpa), min.segment.length = 0, box.padding = 0.5) +
  scale_x_continuous(name = "Change in Total SSB/SSB0", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2), limits = c(0,0.3)) +
  scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) +
  scale_color_manual(values = c("tomato", "steelblue"),name = '') +
  # guides(fill = guide_legend(order = 1, override.aes = list(size = 4))) +
  scale_fill_viridis(name = "% MPA", labels = scales::label_percent(accuracy = 1),guide = guide_colorbar(ticks.colour = "black",frame.colour = "black"), option = "cividis") +
  labs(tag = "C")  +
  theme(
    legend.position = c(0.73, .73),
    legend.background = element_rect(fill = "transparent"),
    legend.text = element_text(size = 10)
  )
coral_fleet_yield <- coral_fleet_results |> 
  mutate(fleet= fct_relabel(fleet, titler)) %>% 
  ggplot(aes(prop_mpa, delta_yield, color = fleet)) +
    geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line() +
  facet_wrap( ~ placement_strategy, labeller = labeller(placement_strategy = titler)) +
  scale_x_continuous(name = "MPA Size",
                     labels = scales::label_percent(accuracy = 1)) +
  scale_y_continuous(name = "Yield",
                     labels = scales::label_percent(accuracy = 1)) + 
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = 8),
        legend.margin = margin())  + 
    scale_color_discrete(name = '') + 
    labs(tag = "A")

fig_4 <- ((coral_fleet_yield / coral_critter_bio) | (coral_frontier + plot_layout(guides = "keep")) ) & theme(strip.text = element_text(size  = 6), plot.margin = margin(.05, .05, .05, .05, "cm"), panel.spacing = unit(0.25, "lines"), legend.text = element_text(size = 8))

 ggsave(
   filename = file.path(results_path, "fig_4.pdf"),
   fig_4,
   width = big_width,
   units = "mm"
 )

fig_4

@@

# blue_frontier / blue_critter_bio + plot_layout(heights = c(1,2))
fig_5 <- (blue_fleet_yield / 
blue_critter_bio) + plot_layout(heights = c(1.5,2))

 ggsave(
   filename = file.path(results_path, "fig_5.pdf"),
   fig_5,
   width = big_width,
   units = "mm"
 )

fig_5
```

### 2023-10-25 04_marlin_paper_si.qmd
first look edits

>>>>

```{r}
#| label: fig-coral-ssb
#| fig-cap: "Spawning biomass divided by unfished spawning biomass in the time period prior to implementation of MPAs for the coral reef case study."

critter_biomass <-
  map_df(coral_sim, ~ map_df(.x, ~tibble(ssb = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
  mutate(step = marlin::clean_steps(step)) %>%
  left_join(grid, by = "patch") %>% 
  group_by(critter, step) %>% 
  summarise(ssb = sum(ssb)) %>% 
  group_by(critter) %>% 
  mutate(depletion = ssb / ssb[step == 0])

fig_s2 <- critter_biomass %>%
  ungroup() %>% 
  filter(step == max(step)) %>% 
  ggplot(aes(reorder(critter, depletion), depletion)) +
  geom_col() +
  scale_y_continuous(limits = c(0, 1), name = "SSB/SSB0", expand = expansion(mult = c(0.0,0.05))) +
  labs(x = "") + 
  coord_flip()

fig_s2
```
